Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{ Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001. }

### 2021-05-04 paper/example.tex
Add figures

\begin{figure*}
\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
\label{fig:example}
\end{figure*}

\figref{fig:example} is an AADL model of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study presented later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight control as the UAV location changes. It is a legacy component that is not subject to change. The alert manager never raises the alert as there is no capacity in the initial implementation to know when an alert is needed.

@@

\begin{figure}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

This work uses an assume guarantee modeling paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE). The contracts constrain the inputs and state properties of the output for component models in the system. In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component. The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See (cite repository) for the full model with contracts.}

@@

The expression for \texttt{yesterday} is helpful to understanding data streams and the synchronous data-flow semantics. It uses the \texttt{->} operator for initialization. The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.  The right operand is the value of \texttt{yesterday} after the initial step. Here the stream nature of expressions is apparent with the \texttt{pre} operator referring to the value of \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response. The \emph{Alert if start is not bounded relative to a request} defines the behavior of the correctly implemented system which is that it alerts if the bounded response policy does not hold. In this way, the contracts model the expected input and output of the system as a whole and between components that implement the system.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). No unexpectedly, since the model of the route planner leaves the output behavior unconstrained, the component can do anything since it is not trusted, the model checker proves that it does not implement the system requirements.

\begin{figure*}
  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
  \label{fig:hardened}
\end{figure*}

\begin{figure}
  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
  \label{fig:assurance}
\end{figure}

The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). The contract model for the filter only passes well-formed output from the route planner (see \figref{fig:assurance}(a)). The monitor only forwards responses that meet the bounded response property (see \figref{fig:assurance}(b)). The contract models for both components are based on the contract in \figref{fig:sw}. The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b). The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.

The high-assurance components are automatically synthesized from the contract models for the final deployed system. Notice that the contracts in \figref{fig:assurance} completely specify the components' behavior stating both the conditions for output events and the values of outputs at those events. Synthesizing a component first turns it into its equivalent CakeML code.
Most expressions have a direct CakeML counterpart. The \texttt{pre} expressions are handled by creating state variables to capture the values in the current step to use in the next step.

>>>>

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.4]{example.png} \\
      (a) \\
      \includegraphics[scale=0.4]{example-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
\caption{Automated UAV route planning system. (a) Unhardened system. (b) Failure certificate.}
\label{fig:example}
\end{figure*}

\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study presented later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight control as the UAV location changes. It is a legacy component that is not subject to change. The alert manager never raises the alert as there is no capacity in the initial design to know when an alert is needed.

@@

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.4]{sw.png}
  \end{center}
  \caption{The SW component contract.}
  \label{fig:sw}
\end{figure}

This work uses an assume guarantee paradigm, in the form of contracts, to model and hierarchically prove properties of the system (cite AGREE). The contracts constrain the inputs and state properties output for component models. In this example, the model for the route planner assumes its input is well-formed, but it makes no guarantees about its output as it is an untrusted third party component. The contract for the legacy waypoint manager makes similar assumptions about its input, but unlike the route planner, the model constrains its output to be well-formed and coincide with the mission command from the route planner and the air vehicle location.\footnote{See (cite repository) for the full model with contracts.}

@@

The expression for \texttt{yesterday} is helpful to understanding data streams and the synchronous data-flow semantics. It uses the \texttt{->} operator for initialization. The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{false} because there is no yesterday at the start.  The right operand is the value of \texttt{yesterday} after the initial step. Here the stream nature of expressions is apparent with the \texttt{pre} operator referring to the value of \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response. The \emph{Alert if start is not bounded relative to a request} defines the behavior of the correctly implemented system which is that it alerts if the bounded response policy does not hold. In this way, the contracts model the expected input and output of the system as a whole and also between components that implement the system.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). Not unexpectedly, since the model of the route planner leaves the output behavior unconstrained (i.e., the component can do anything since it is not trusted) the model checker proves that the component composition does not implement the system requirements.

\begin{figure*}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{hardened.png} \\
      (a) \\
      \includegraphics[scale=0.4]{hardened-certificate.png} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{Hardened UAV system. (a) The implementation with high-assurance components. (b) Passing certificate.}
  \label{fig:hardened}
\end{figure*}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.45]{filter.png} \\
      (a) \\
      \includegraphics[scale=0.4]{monitor.png} \\
      (b)
    \end{tabular}
  \end{center}
  \caption{High-assurance component contracts. (a) The filter. (b) The monitor.}
  \label{fig:assurance}
\end{figure}

The system implementation is cyber-hardened by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}(a). A filter takes a stream of data and delivers a new stream of data where some property is enforced. Its stylized form generated by the CASE tools is shown in \figref{fig:assurance}(a). The user must provide the meaning of the property, a predicate, while the tool generates everything else. In this example, the property is the assumption made by the waypoint manager about the automation response being well-formed. The guarantee is such that it has a trivial direct mapping to synthesized code by completely specifying the behavior of the output.

A monitor captures a relation over time on data and is able to reason about temporal properties. In this example, the monitor only forwards responses that meet the bounded response property as shown in \figref{fig:assurance}(b). Like the filter, the generated monitor, with its contract, is stylized for synthesis only requiring the user to indicate the policy and some extra information for the output. The policy is taken from the contract in \figref{fig:sw} since it is the monitor that needs to alert violations of the bounded response.

The proof certificate for the new cyber-hardened implementation is in \figref{fig:hardened}(b). The high-assurance components guarantee the correct behavior of the SW implementation in the presence of the considered cyber attacks.

The high-assurance components are automatically synthesized from the contract models for the final deployed system. Synthesis assumes the existence of the communication fabric and a scheduler that implements the synchronous data-flow semantics by dispatching components in dependency order. For this example, and the later case study, the HAMR framework provides both requirements.

\newsavebox{\boxa}
\begin{lrbox}{\boxa}
\begin{lstlisting}
fun filter_step () =
let val () = Utils.clear_buf filterBuffer
    val () = API.callFFI "get_input" "" filterBuffer
    val string = Utils.buf2string filterBuffer
in
    if WELL_FORMED_AUTOMATION_RESPONSE string
    then
        let val string = Word8Array.substring filterBuffer 1
                         (Word8Array.length filterBuffer - 1)
        in API.callFFI "put_output" string Utils.emptybuf
        end
    else print"Filter rejects message.\n"
end
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\boxa}}
    \end{tabular}
  \end{center}
  \caption{Synthesized CakeML for the filter.}
  \label{fig:cakeml}
\end{figure}

Synthesis targets CakeML, and CakeML provides verified compilation to binaries for several different platforms. The generated code for the filter is shown in \figref{fig:cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate. The synthesis is proven to exactly implement the meaning of the contract in the synchronous data-flow semantics in the CakeML assuming perfect communication and scheduling.

### 2021-05-04 paper/paper.tex
Add example description.
Add figures

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{comment}
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\begin{document}
\title{
  Synthesizing Verified Components for Cyber Assured Systems Engineering 
  \thanks{
    Funded by DARPA Cyber Assured Systems Engineering (CASE)
  }
}

@@

% \section{Simple Example}
% \input{example}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{comment}
\usepackage{paralist}
\usepackage{graphicx}
\graphicspath{{./figs/}} 
\usepackage{listings}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.88}

\lstset{frame=none,
  basicstyle=\ttfamily,
  language=ML,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  numbers=none,
  numberstyle=\tiny\color{gray},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=false,
  breakatwhitespace=true,
  tabsize=2,
  morekeywords={function, method, then, reads, modifies, requires, ensures, var, assume, assert}
}

\hyphenation{op-tical net-works semi-conduc-tor}

@@

\begin{document}
\title{
  Synthesizing Verified Components for Cyber Assured Systems Engineering 
  \thanks{
    Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001.
  }
}

@@

\section{Simple Example}
\input{example}
